---
title: "Statistical Uncertainty in the Ranking of Journals and Universities"
subtitle: "Replication of Mogstad, Romano, Shaikh & Wilhelm (2022)"
author: "Jason Jankoski, Alan Lamb, Linna Wang"
date: "ECON 687 — University of Maryland"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE
)
knitr::opts_knit$set(root.dir = "../")
```

```{r load-packages, include=FALSE}
library(ggplot2)
library(dplyr)
library(readxl)
library(csranks)
```


# The Paper

## What Question Does It Ask?

Rankings of journals and universities are used constantly in academic hiring,
promotion, and funding decisions. But they are almost always reported as a
single number — journal X is ranked 5th — with no indication of how precisely
that rank is estimated.

This paper asks: **given that the underlying measures (like impact factors) are
themselves estimated with uncertainty, how much uncertainty surrounds the
resulting ranks?**

The authors develop confidence sets for ranks: rather than reporting a single
rank, they report the range of ranks that cannot be statistically rejected at
the 95% confidence level.

## Method in Plain Terms

The core idea is straightforward. An impact factor is an average — it is
estimated from a finite number of articles and carries a standard error. When
you rank things by noisy estimates, the resulting ranks are also noisy.

The authors' method:

- Takes the point estimates (impact factors) and their standard errors as inputs
- Uses a bootstrap procedure to simulate the distribution of possible rankings
- Constructs a confidence set for each entity's rank — the set of ranks that
  are statistically consistent with the data at the 95% level

The method is implemented in the R package `csranks`, written by the authors
and available on GitHub.

## What Did They Find?

Two empirical applications:

**Journals** — ranking 30 economics journals by 2011 impact factor:

- 10 of the 30 journals cannot be statistically distinguished from the top 5
- Only 1 journal can be ruled out of the bottom 25

**Universities** — ranking 100 universities using publication records from
Research Papers in Economics (RePEC):

- The top and bottom of the ranking are fairly precise
- Substantial uncertainty throughout the middle


# The Data

## Two Datasets — Both Included in Replication Package

No external data access required. Both files are in the `data/` folder.

| Dataset | Source | File |
|---|---|---|
| Journal impact factors | Stern (2013) | `data/journal IF/JEL_Data.xlsx` |
| University publications | Zimmermann (2013) | `data/Repec/top100unis.csv` |

The journal data was assembled and shared by David Stern. The university data
comes from RePEC (Research Papers in Economics), a self-reported database of
economists' affiliations and publications maintained by Christian Zimmermann.

## Journal Data — Loading and Structure

The Excel file has one sheet per journal with raw article-level citation counts.
The authors' `convertdata.R` script processes this into impact factors and
standard errors.

```{r load-journal-data}
# Load the processed journal IF data
load(file.path("data", "journal IF", "IFdata.RData"))

# The exact 30-journal list from Kalaitzidakis, Mamuneas & Stengos (2003)
top30names <- c("JEL", "QJE", "JEP", "JFE", "JPE", "Econometrica",
                "AER", "REStud", "RESTAT", "JLE", "JEEM", "JHumRes",
                "J INT ECON", "EconJ", "JMonEcon", "JEconometrics",
                "RAND J ECON", "J BUS ECON STAT", "JPubEcon",
                "JAppliedEconmet", "EUR ECON REV", "IER", "JET",
                "OXFORD B ECON STAT", "GAME ECON BEHAV",
                "J ECON DYN CONTROL", "SCAND J ECON",
                "ECONOMET THEOR", "Econ Theory", "EL")

# Filter to top 30 and rank by impact factor
jIF <- journalIF %>%
  filter(!is.na(IF), IF > 0, journal %in% top30names) %>%
  arrange(desc(IF)) %>%
  mutate(rank = row_number())

cat("Journals matched:", nrow(jIF), "of 30")
```

## Journal Data — What We're Working With

```{r journal-table}
jIF %>%
  select(Rank = rank, Journal = journal, IF, SE, N_articles = n) %>%
  mutate(IF = round(IF, 3), SE = round(SE, 3)) %>%
  head(10) %>%
  knitr::kable(caption = "Top 10 journals by 2011 impact factor")
```

## Journal Data — Descriptive Statistics

```{r journal-summary}
cat("--- All 231 Valid Journals ---\n")
jall <- journalIF %>% filter(!is.na(IF), IF > 0)
cat("Mean IF:  ", round(mean(jall$IF), 4), "\n")
cat("Median IF:", round(median(jall$IF), 4), "\n")
cat("SD of IF: ", round(sd(jall$IF), 4), "\n")
cat("Range:    [", round(min(jall$IF), 4), ",", round(max(jall$IF), 4), "]\n\n")

cat("--- KMS Top 30 Journals ---\n")
cat("Mean IF:  ", round(mean(jIF$IF), 4), "\n")
cat("Median IF:", round(median(jIF$IF), 4), "\n")
cat("SD of IF: ", round(sd(jIF$IF), 4), "\n")
cat("Mean SE:  ", round(mean(jIF$SE), 4), "\n")
cat("Range:    [", round(min(jIF$IF), 4), ",", round(max(jIF$IF), 4), "]\n")
```

## University Data — Loading and Structure

```{r load-university-data}
# Load the cleaned top-100 university dataset
unis <- read.csv(file.path("data", "Repec", "top100unis.csv")) %>%
  filter(!is.na(IFmean)) %>%
  arrange(desc(IFmean)) %>%
  mutate(rank = row_number())

# Column definitions:
#   name   = university name
#   n      = number of affiliated authors in RePEC
#   IFmean = average impact factor of the university's publications
#   IFse   = standard error of that average
#   CTmean = average citation count
#   CTse   = standard error of citation mean

glimpse(unis)
```

## University Data — Descriptive Statistics

```{r university-summary}
cat("--- 100 Universities ---\n")
cat("Mean IFmean:   ", round(mean(unis$IFmean), 4), "\n")
cat("Median IFmean: ", round(median(unis$IFmean), 4), "\n")
cat("SD of IFmean:  ", round(sd(unis$IFmean), 4), "\n")
cat("Mean IFse:     ", round(mean(unis$IFse), 4), "\n")
cat("Total authors: ", sum(unis$n), "\n")
cat("IF range:      [", round(min(unis$IFmean), 4), ",",
                        round(max(unis$IFmean), 4), "]\n")
```

## University Data — Top 20

```{r university-table}
unis %>%
  slice(1:20) %>%
  select(Rank = rank, University = name, IFmean, IFse, N_authors = n) %>%
  mutate(IFmean = round(IFmean, 3), IFse = round(IFse, 3)) %>%
  knitr::kable(caption = "Top 20 universities by average impact factor")
```


# The Code

## Overall Pipeline

The replication follows the authors' code structure exactly:

```
data/journal IF/convertdata.R    # cleans Excel → IFdata.RData
data/journal IF/journalranking.R # computes confidence sets → Figure 1
data/Repec/cleandata.R           # cleans RePEC raw data
data/Repec/keeptop100.R          # filters to top 100 universities
data/Repec/emp.topunis.R         # computes confidence sets → Figure 2
```

Our `code/00_data_check.R` runs the key steps independently so we can confirm
the pipeline works without running the full `master.R`.

## Data Import — What This Code Does

The journal data comes in as a messy Excel file. Each of the 233 sheets is one
journal, with rows representing individual articles and columns representing
years. There is no single flat table to load directly.

```{r show-sheets, eval=FALSE}
# This is what the raw Excel looks like before convertdata.R runs
sheet_names <- excel_sheets(file.path("data", "journal IF", "JEL_Data.xlsx"))
cat("Number of sheets:", length(sheet_names))
cat("First 5 sheets:", head(sheet_names, 5))

# Each sheet looks like this:
# rows    = individual articles published in that journal
# columns = years (how many citations that article received each year)
# bottom  = summary stats: mean, SD, SE of the mean
```

`convertdata.R` loops through every sheet, extracts the mean and standard error
from the bottom rows, and assembles a single clean data frame.

## Data Wrangling — Filtering to the KMS Top 30

The full dataset has 232 journals. The paper focuses on 30 specific journals
pre-selected by Kalaitzidakis, Mamuneas & Stengos (2003) — an earlier study
that identified leading outlets in economics based on citation data from
1994–1998. We filter to exactly those 30 by name.

```{r show-filter, eval=FALSE}
jIF <- journalIF %>%
  filter(
    !is.na(IF),              # remove journals with no impact factor data
    IF > 0,                  # remove journals with zero citations
    journal %in% top30names  # keep only the KMS 30
  ) %>%
  arrange(desc(IF)) %>%      # sort highest IF first
  mutate(rank = row_number()) # assign point ranks 1 through 30
```

The `arrange` and `mutate` steps give us the point estimate of each journal's
rank — the rank you would assign if you ignored uncertainty entirely.

## The Key Statistical Step — csranks

This is the core of the replication. The `csranks()` function takes the point
estimates and their uncertainty, runs a bootstrap simulation, and returns
confidence sets for ranks.

```{r show-csranks, eval=FALSE}
# Convert SE vector to a diagonal covariance matrix
# (csranks requires a matrix, not a vector)
# diag(SE^2) places variances on the diagonal, zeros off-diagonal
# This assumes estimation errors are independent across journals
Sigma <- diag(jIF$SE^2)

CS <- csranks::csranks(
  jIF$IF,          # point estimates (impact factors)
  Sigma = Sigma,   # covariance matrix of estimation errors
  coverage = 0.95, # 95% confidence level
  stepdown = TRUE, # stepdown procedure (more powerful than single-step)
  simul    = FALSE, # marginal (not simultaneous) confidence sets
  R        = 1000, # number of bootstrap draws
  seed     = 101   # for reproducibility
)

# Output: CS$L = lower bound of rank confidence set
#         CS$U = upper bound of rank confidence set
jIF$CSL <- CS$L
jIF$CSU <- CS$U
```

## What stepdown and simul Mean

Two parameters worth understanding:

**`stepdown = TRUE`** — uses a stepwise testing procedure. Instead of testing
all ranks simultaneously at once, it iterates, removing journals whose ranks
are already determined. This gives tighter (more informative) confidence sets
than a single-step approach.

**`simul = FALSE`** — produces *marginal* confidence sets. Each journal's
confidence set covers its true rank with 95% probability individually.
When `simul = TRUE`, the sets are *simultaneous* — all true ranks are covered
jointly with 95% probability. Simultaneous sets are wider. The paper shows both;
Figure 1 uses marginal sets.


# Figures

## Figure 1 — Journal Rankings (Proof of Concept)

```{r figure1, fig.height=4.5, fig.width=9, echo=FALSE}
# Load pre-computed confidence sets if available
# (running bootstrap live takes ~2 minutes)
cs_path <- file.path("data", "journal IF", "journalIFrankings.RData")
if (file.exists(cs_path)) {
  load(cs_path)
  plot_data <- journalIF.top30
} else {
  Sigma <- diag(jIF$SE^2)
  CS <- csranks::csranks(jIF$IF, Sigma=Sigma, coverage=0.95,
                          stepdown=TRUE, simul=FALSE, R=1000, seed=101)
  plot_data <- jIF %>% mutate(margCSL = CS$L, margCSU = CS$U, rank = rank)
}

ggplot(plot_data, aes(x = rank, y = reorder(journal, rank))) +
  theme_bw() +
  geom_point(size = 2) +
  geom_errorbar(aes(xmin = margCSL, xmax = margCSU), width = 0.4) +
  scale_x_continuous(name = "Rank", limits = c(1,30),
                     breaks = c(1, 5, 10, 15, 20, 25, 30)) +
  ylab(NULL) +
  labs(title = "Ranking of 30 Journals by 2011 Impact Factor",
       subtitle = "95% marginal confidence sets for ranks")
```

## Figure 2 — University Rankings (Proof of Concept)

```{r figure2, fig.height=4.5, fig.width=9, echo=FALSE}
Sigma_u <- diag(unis$IFse^2)
CS_u <- csranks::csranks(unis$IFmean, Sigma=Sigma_u, coverage=0.95,
                          stepdown=TRUE, simul=FALSE, R=1000, seed=101)
unis$margCSL <- CS_u$L
unis$margCSU <- CS_u$U

ggplot(unis, aes(x = rank, y = reorder(name, rank))) +
  theme_bw() +
  geom_point(size = 0.7) +
  geom_errorbar(aes(xmin = margCSL, xmax = margCSU), width = 0.4) +
  scale_x_continuous(name = "Rank", limits = c(1,100),
                     breaks = c(1, seq(10,100,by=10))) +
  ylab(NULL) +
  labs(title = "Ranking of 100 Universities by Impact Factor",
       subtitle = "95% marginal confidence sets for ranks") +
  theme(axis.text.y = element_text(size = 5))
```


# Next Steps

## What Remains

**For the final presentation (April 30):**

- Run `master.R` fully to reproduce all figures exactly as they appear in the paper
- Compare our replicated figures side by side with the paper's originals
- Discuss where replication succeeded and where any differences appeared
- Deeper walkthrough of the econometric foundations of the confidence set method

**Remaining code to cover:**

- `data/Repec/cleandata.R` — how the raw RePEC text file is processed
- `data/Repec/keeptop100.R` — how the top 100 universities are selected
- `data/journal IF/journalranking.R` — the full ranking script with all appendix figures
