---
title: "Statistical Uncertainty in the Ranking of Journals and Universities"
subtitle: "Replication of Mogstad, Romano, Shaikh & Wilhelm (2022)"
author: "Jason Jankoski, Alan Lamb, Linna Wang"
date: "ECON 687 — University of Maryland"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
    css: custom.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE
)
knitr::opts_knit$set(root.dir = "../")
```

```{r install-packages, include=FALSE}
if (!requireNamespace("ggplot2",  quietly = TRUE)) install.packages("ggplot2")
if (!requireNamespace("dplyr",    quietly = TRUE)) install.packages("dplyr")
if (!requireNamespace("readxl",   quietly = TRUE)) install.packages("readxl")
if (!requireNamespace("devtools", quietly = TRUE)) install.packages("devtools")
if (!requireNamespace("csranks",  quietly = TRUE))
  devtools::install_github("danielwilhelm/R-CS-ranks")
```

```{r load-packages, include=FALSE}
library(ggplot2)
library(dplyr)
library(readxl)
library(csranks)
```


# The Paper

## What Question Does It Ask?

Rankings of journals and universities drive real decisions in hiring, promotion,
and funding. They are almost always reported as a single number with no
indication of how precisely that rank is estimated.

This paper asks: **given that impact factors are themselves estimated with
uncertainty, how much uncertainty surrounds the resulting ranks?**

The authors construct confidence sets for ranks. Rather than saying a journal
is ranked 5th, they report the full range of ranks that cannot be statistically
rejected at the 95% level.

## Method in Plain Terms

An impact factor is an average computed from a finite number of articles. It
carries a standard error. When you rank things by noisy estimates, the ranks
themselves are noisy.

The pipeline:

1. Start with point estimates (impact factors) and their standard errors
2. Build a covariance matrix capturing estimation uncertainty
3. Run a bootstrap simulation to generate possible rankings under that uncertainty
4. Report, for each journal or university, the set of ranks consistent with the data

The method is implemented in `csranks`, an R package written by the authors.

## What Did They Find?

**Journals** — 30 economics journals ranked by 2011 impact factor:

- Any of 10 different journals could plausibly be in the top 5
- The middle of the ranking has confidence bars 8 to 12 ranks wide
- At the bottom, only the last few journals are locked in place

**Universities** — 100 universities ranked by average publication impact factor:

- Rankings at the very top and bottom are fairly precise
- Substantial uncertainty throughout the middle of the distribution

The main takeaway: point rankings overstate what the data can actually tell us.


# The Data

## Journal Data — What the Raw Excel Looks Like

One sheet per journal, one row per article. Columns are years of citations received.

```
| Avg/yr | ... | 2006 | 2007 | 2008 | 2009 | 2010 | 2011 |
|--------|-----|------|------|------|------|------|------|
| 4.14   | ... | 2    | 1    | 8    | 3    | 7    | 6    |
| 9.57   | ... | 8    | 8    | 8    | 10   | 18   | 12   |
| 2.57   | ... | 0    | 4    | 6    | 3    | 1    | 3    |
| 14.71  | ... | 0    | 3    | 12   | 23   | 24   | 31   |
```

The impact factor is the average citations received in 2011 per article.
`convertdata.R` loops through all 233 sheets and computes mean and SE for each.

## Journal Data — convertdata.R

```{r show-convertdata, eval=FALSE}
for (s in sheetnames[-1]) {
  X <- as.data.frame(mylist[s])             # load one journal's sheet
  X <- X[-((nrow(X)-3):nrow(X)),]           # drop the summary rows at the bottom
  col2011 <- names(X)[lastcharscn==".2011"] # find the 2011 column by name

  journalIF <- rbind(journalIF,             # stack results row by row
    c(mean(X[, col2011]),                   # impact factor = mean citations in 2011
      sd(X[, col2011]) / sqrt(nrow(X)),     # SE = sd / sqrt(n articles)
      sd(X[, col2011]),                     # sigma = standard deviation
      nrow(X)))                             # n = number of articles
}
```

## Journal Data — Filtering to the KMS 30

The paper focuses on 30 journals pre-selected by Kalaitzidakis, Mamuneas &
Stengos (2003), an earlier study identifying leading outlets in economics.
We filter the 232-journal dataset to exactly those 30 by name.

```{r load-journal-data}
load(file.path("data", "journal IF", "IFdata.RData"))

top30names <- c("JEL", "QJE", "JEP", "JFE", "JPE", "Econometrica",
                "AER", "REStud", "RESTAT", "JLE", "JEEM", "JHumRes",
                "J INT ECON", "EconJ", "JMonEcon", "JEconometrics",
                "RAND J ECON", "J BUS ECON STAT", "JPubEcon",
                "JAppliedEconmet", "EUR ECON REV", "IER", "JET",
                "OXFORD B ECON STAT", "GAME ECON BEHAV",
                "J ECON DYN CONTROL", "SCAND J ECON",
                "ECONOMET THEOR", "Econ Theory", "EL")

jIF <- journalIF %>%
  filter(!is.na(IF), IF > 0, journal %in% top30names) %>%
  arrange(desc(IF)) %>%
  mutate(rank = row_number())

cat("Journals matched:", nrow(jIF), "of 30")
```

## Journal Data — Usable form

```{r journal-table, echo=FALSE}
jIF %>%
  select(Rank = rank, Journal = journal, IF, SE, N_articles = n) %>%
  mutate(IF = round(IF, 3), SE = round(SE, 3)) %>%
  head(10) %>%
  knitr::kable(caption = "Top 10 journals by 2011 impact factor")
```

## RePEC Data — What the Raw File Looks Like

One row per author per publication type. The affiliation field encodes
institution code and weight. Columns are article-level impact factors or citation counts.

```
affiliation                             author  type  X1      X2      X3
RePEc:edi:aabasch+1                     1       IF    8.8677  9.5333  1.7667
RePEc:edi:aabasch+1                     1       CT    17      1       1
RePEc:edi:aabasch+*RePEc:edi:faphijo+*  2       IF    0.3513  0.9251  0.3513
```

- `+1` = full weight assigned to this institution
- `+*` = weight unspecified, author has multiple affiliations
- Each author has two rows — one IF row and one CT row

`cleandata.R` parses the affiliation strings, computes weighted averages per
institution, aggregates across all authors, and removes non-universities.
`keeptop100.R` filters to the top 100 and saves `top100unis.csv`.

## University Data — Loading and Wrangling

The university data required less wrangling than the journal data. By the time
we load `top100unis.csv` the authors' cleaning scripts have already computed
average impact factors and standard errors per university.

```{r load-university-data}
unis <- read.csv(file.path("data", "Repec", "top100unis.csv")) %>%
  filter(!is.na(IFmean)) %>%       # remove any missing values
  arrange(desc(IFmean)) %>%         # sort highest impact factor first
  mutate(rank = row_number())       # assign point ranks 1 through 100

# IFmean = average IF of publications by authors affiliated with that university
# IFse   = standard error of that average
# n      = number of affiliated authors in RePEC

cat("Universities:", nrow(unis), "| Total authors:", sum(unis$n))
```

## University Data — Usable form

```{r university-table, echo=FALSE}
unis %>%
  slice(1:10) %>%
  select(Rank = rank, University = name, IFmean, IFse, N_authors = n) %>%
  mutate(IFmean = round(IFmean, 3), IFse = round(IFse, 3)) %>%
  knitr::kable(caption = "Top 10 universities by average impact factor")
```


# Descriptive Statistics and Methods
<!-- Linna: feel free to rename this section header if something fits better -->

## The Core Econometric Problem — Defining "True Rank"

* **The Problem:** We observe estimated impact factors, not the "true" population impact factors.  
* Therefore, any ranking is subject to statistical uncertainty. 
* **To define a rigorous rank, we must clear a fundamental statistical hurdle:** \

$$r_i = 1 + \sum_{j \neq i} \mathbb{I}(\theta_j > \theta_i)$$

| **Parameter $(\theta)$:** | **Indicator $\mathbb{I}(\cdot)$:** | **The Rank $(r_i)$:** |
| :--- | :--- | :--- |
| The **true**, unobservable quality score of University $i$ and $j$. This is the noise-free value we wish we knew. | A binary variable. It returns **1** if $\theta_j > \theta_i$ (the "better-than" test) and **0** otherwise. | The sum counts how many schools $j$ are better than $i$. The "+1" ensures the top school is Rank 1. |


<!-- How do we define a 'True Rank'? If we denote the true performance of an institution i as θi,its true rank is simply 1 plus the number of other institutions that have a strictly better performance. -->

<!-- To see how this works in practice, let's use a local example. Imagine we are comparing UMD alongside another university right here in the DMV, like George Mason. UMD is our institution i. The indicator function here—the math inside the summation—asks a simple yes-or-no question for every other university j: 'Is your true score strictly higher than UMD's?' -->

<!-- If George Mason does have a strictly higher true score, the indicator returns a 1, which gets added to the sum, pushing UMD's rank down by one spot. If UMD is higher, it returns a 0. We do this for every school in our dataset. If exactly 14 schools trigger that '1', then UMD's rank is 1 plus 14, placing it at rank 15. -->

<!-- The formula is perfectly straightforward. However, because we don't have the luxury of knowing the true θ—we only have our statistical estimates—our calculated ranks are inherently uncertain. We need a way to quantify that uncertainty when schools like UMD and George Mason are statistically tied. -->

## Journal Data - Summary Statistics

<div class="columns-2" style="height: 1000px;">

```{r journal-summary, echo=FALSE}
jall <- journalIF %>% filter(!is.na(IF), IF > 0)

journal_stats <- data.frame(
  Statistic = c("Mean IF", "Median IF", "SD of IF", "Mean SE", "Min IF", "Max IF"),
  `All 231 Journals` = c(
    round(mean(jall$IF), 3),
    round(median(jall$IF), 3),
    round(sd(jall$IF), 3),
    NA,
    round(min(jall$IF), 3),
    round(max(jall$IF), 3)
  ),
  `KMS Top 30` = c(
    round(mean(jIF$IF), 3),
    round(median(jIF$IF), 3),
    round(sd(jIF$IF), 3),
    round(mean(jIF$SE), 3),
    round(min(jIF$IF), 3),
    round(max(jIF$IF), 3)
  ),
  check.names = FALSE
)

knitr::kable(journal_stats)
```
<div style="break-before: column;">

**Reading the table**

- **IF** is the 2011 impact factor, the mean citations per 2008 article
- **SE** is the standard error of that mean
- The KMS 30 sample has a higher mean IF than the full set, by construction
- Substantial variance across the metrics, the noise that makes ranks hard to assign

</div>

</div>

<!-- Looking at the summary statistics for our journal dataset, the key takeaway is the substantial variance present across the performance metrics. It is exactly this underlying statistical noise that makes assigning a definitive, ordinal rank based purely on point estimates so problematic. -->

## University Descriptive Statistics

<div class="columns-2" style="height: 1000px;">

```{r university-summary, echo=FALSE}
uni_stats <- data.frame(
  Statistic = c("Mean IFmean", "Median IFmean", "SD of IFmean",
                "Mean IFse", "Total authors", "Min IFmean", "Max IFmean"),
  Value = c(
    round(mean(unis$IFmean), 3),
    round(median(unis$IFmean), 3),
    round(sd(unis$IFmean), 3),
    round(mean(unis$IFse), 3),
    sum(unis$n),
    round(min(unis$IFmean), 3),
    round(max(unis$IFmean), 3)
  )
)

knitr::kable(uni_stats)
```
<div style="break-before: column;">

**Reading the table**

- **IFmean** is each university's average article-level impact factor, weighted by author affiliation
- **IFse** is the standard error of that average
- The 100 universities are tightly packed around a median IFmean near 10
- The SEs are large enough to blur the boundaries between adjacent ranks

</div>

</div>

<!-- As you can see here, our 100 universities are tightly packed around a median score of 10, but the standard errors at the bottom are large enough to blur the lines between them. Just eye-balling this, we can tell there is significant statistical overlap. It’s impossible to just sort these scores 1 through 100. So, how do we actually handle this overlap rigorously? Before I show you the specific formulas, let's look at the overarching strategy we use to process this.-->

## The Full Pipeline

```
data/journal IF/convertdata.R     # Excel (233 sheets) -> IFdata.RData
data/journal IF/journalranking.R  # confidence sets -> Figure 1
data/Repec/cleandata.R            # raw RePEC text -> cleaned data
data/Repec/keeptop100.R           # filter to top 100 universities
data/Repec/emp.topunis.R          # confidence sets -> Figure 2
```
The four-step logic repeated for both datasets:

> **Load clean data, filter to sample, run bootstrap, plot results**

<!-- To solve this, we built a 4-step pipeline that feeds these noisy impact factors into the authors' csranks package. We supply the covariance matrix, run a step-down bootstrap with 1,000 draws, and output the upper and lower rank bounds—which Jason will show you visually later. -->


## The Deep Dive: Confidence Sets

To define a rigorous rank, we must clear a fundamental statistical hurdle. For any two institutions $i$ and $j$, we test if their estimated performance gap is distinct from the statistical noise:

$$Z_{i,j} = \left| \frac{\widehat{\theta}_j - \widehat{\theta}_i}{SE_{i,j}} \right| > cv$$

We face a critical choice in how we set that critical value ($cv$):

| Marginal Confidence Sets | Simultaneous Confidence Sets |
| :--- | :--- |
| **The Choice:** Use the standard critical value (e.g., $cv \approx 1.96$). | **The Choice:** Use an adjusted $cv$ calculated for joint estimation. |
| **Posing:** Answers: *"Does this single university belong in the Top 10?"* | **Posing:** Answers: *"Are all these ranks accurate relative to one another with 95% confidence?"* |
| **Danger:** High "False Discovery Rate" when ranking many institutions at once. | **Required:** Essential for rigorous, full-dataset rank reporting. |

<!-- Transition and The Statistical Hurdle: -->
<!-- Now that we’ve seen the Full Pipeline, you’ll notice that the most critical stage happens at the very end: moving from raw estimates to defensible ranks. To understand how we do that, we need to look at the specific statistical hurdle we have to clear. To determine if one institution is truly better than another, we use the test statistic shown at the top of the slide. We calculate the difference between two estimates, divide by their joint standard error, and check if that value clears a specific critical value, or cv. -->

<!-- Comparing Marginal and Simultaneous Sets: -->
<!-- The choice of that cv is where the econometric rigor of this paper lives. On the left, we have Marginal Confidence Sets, which use the standard critical value of approximately 1.96. While this is fine for comparing two schools in a vacuum, using it for a full list of 100 schools leads to a massive False Discovery Rate. By pure statistical chance, you would claim ranks are different when they are actually just noise. On the right, we have our preferred method: Simultaneous Confidence Sets. Here, we adjust the critical value to account for the entire distribution at once, ensuring we don't overstate our certainty. -->

<!-- Closing and Next Step: -->
<!-- This adjustment is the only way to state, with 95% confidence, that the entire list of rankings is accurate relative to one another. It transforms our results from a simple list into a statistically robust model that accounts for the inherent uncertainty in our data. So, let’s see how this math translates into the final R implementation using the csranks package. -->


## The Key Statistical Step — csranks

```{r show-csranks, eval=FALSE}
# Step 1: build a covariance matrix from the standard errors
# diag(SE^2) puts variances on the diagonal, zeros off-diagonal
# This assumes estimation errors are independent across journals
Sigma <- diag(jIF$SE^2)

# Step 2: run the bootstrap
CS <- csranks::csranks(
  jIF$IF,           # point estimates
  Sigma = Sigma,    # covariance matrix
  coverage = 0.95,  # 95% confidence level
  stepdown = TRUE,  # iterative procedure, gives tighter sets
  simul    = FALSE, # marginal sets (each journal covered individually)
  R        = 1000,  # bootstrap draws
  seed     = 101    # reproducibility
)

# Step 3: attach bounds to the data frame
jIF$CSL <- CS$L   # lower bound of rank confidence set
jIF$CSU <- CS$U   # upper bound of rank confidence set
```

<!-- Now, how do we actually put all this math into practice? We use an R package called csranks, which is specifically built to handle these kinds of complex rankings. This function is essentially the 'engine' of our project—it takes our raw data and standard errors and does the heavy lifting to find the true ranking windows. -->

<!-- If you look at the code on the screen, keep an eye on the line that says simul = FALSE. This is the default starting point for most researchers. It produces what we call Marginal Confidence Sets. Essentially, the software is looking at each journal one by one, almost as if they exist in a vacuum. It calculates a 95% certainty window for each specific journal without worrying about how it relates to the rest of the group. It's a useful first look, but it doesn't give us the full picture of the uncertainty across the entire list. -->

<!-- In our replication, the real 'game-changer' is when we flip that switch to TRUE. That one tiny change tells the software to use the tougher statistical hurdle we just talked about. It shifts the focus from individual journals to the entire system at once. By showing both versions, we can see exactly how our confidence holds up when we move from a narrow view to a rigorous, collective one. With that setup in mind, I'll pass it over to my teammate to show you what these rankings actually look like in the final plots. -->

# The Figures

## From Bootstrap Output to Figure

Once `csranks()` returns the bounds, the figure follows directly.
The bounds become the error bar limits in ggplot.

```{r show-plot, eval=FALSE}
ggplot(jIF, aes(x = rank, y = reorder(journal, rank))) +
  theme_bw() +
  geom_point(size = 2) +             # dot = point rank estimate
  geom_errorbar(                      # bar = confidence set for rank
    aes(xmin = CSL, xmax = CSU),
    width = 0.4
  ) +
  scale_x_continuous(
    name = "Rank",
    limits = c(1, 32),
    breaks = c(1, 5, 10, 15, 20, 25, 30)
  ) +
  ylab(NULL)
```
## Figure 1 — Journal Rankings

```{r figure1, fig.height=5.5, fig.width=9, echo=FALSE}
cs_path <- file.path("data", "journal IF", "journalIFrankings.RData")
if (file.exists(cs_path)) {
  load(cs_path)
  plot_data <- journalIF.top30
} else {
  Sigma <- diag(jIF$SE^2)
  CS <- csranks::csranks(jIF$IF, Sigma=Sigma, coverage=0.95,
                          stepdown=TRUE, simul=FALSE, R=1000, seed=101)
  plot_data <- jIF %>% mutate(margCSL = CS$L, margCSU = CS$U)
}

ggplot(plot_data, aes(x = rank, y = reorder(journal, rank))) +
  theme_bw() +
  geom_point(size = 2) +
  geom_errorbar(aes(xmin = margCSL, xmax = margCSU), width = 0.4) +
  scale_x_continuous(name = "Rank", limits = c(1,32),
                     breaks = c(1, 5, 10, 15, 20, 25, 30)) +
  ylab(NULL) +
  labs(title = "Figure 1: Ranking of 30 Journals by 2011 Impact Factor",
       subtitle = "95% marginal confidence sets for ranks")
```
The x-axis is rank, not impact factor. The horizontal bars show the range of
ranks the data cannot rule out for each journal at the 95% level.

## Figure 1 — Journal Rankings (alt-view)

```{r figure1b, fig.height=4.5, fig.width=9, echo=FALSE}
cs_path <- file.path("data", "journal IF", "journalIFrankings.RData")
if (file.exists(cs_path)) {
  load(cs_path)
  plot_data <- journalIF.top30
} else {
  Sigma <- diag(jIF$SE^2)
  CS <- csranks::csranks(jIF$IF, Sigma=Sigma, coverage=0.95,
                          stepdown=TRUE, simul=FALSE, R=1000, seed=101)
  plot_data <- jIF %>% mutate(margCSL = CS$L, margCSU = CS$U)
}

plot_data$span <- plot_data$margCSU - plot_data$margCSL + 1
plot_data$span_cat <- cut(plot_data$span,
                          breaks = c(-Inf, 5, 10, 20, Inf),
                          labels = c("<=5","6-10","11-20",">20"))

ggplot(plot_data, aes(x = rank, y = reorder(journal, rank))) +
  theme_bw() +
  geom_errorbar(aes(xmin = margCSL, xmax = margCSU, colour = span_cat),
                width = 0.4, size = 0.8, show.legend = TRUE) +
  geom_point(aes(colour = span_cat), size = 2, show.legend = TRUE) +
  scale_colour_manual(name = "Span (ranks)",
                      values = c("<=5"   = "red",
                                 "6-10"  = "black",
                                 "11-20" = "orange",
                                 ">20"   = "blue")) +
  scale_x_continuous(name = "Rank", limits = c(1,32),
                     breaks = c(1, 5, 10, 15, 20, 25, 30)) +
  ylab(NULL) +
  labs(title = "Ranking of 30 Journals by 2011 Impact Factor",
       subtitle = "95% marginal confidence sets for ranks") +
  theme(axis.text.y = element_text(size = 8),
        legend.position = "right")
```


## Figure 2 — University Rankings Code

Same structure as Figure 1 but applied to 100 universities. The csranks
bootstrap runs first, then the bounds feed directly into ggplot.

```{r show-plot2, eval=FALSE}
# run bootstrap for university ranks
Sigma_u <- diag(unis$IFse^2)
CS_u <- csranks::csranks(unis$IFmean, Sigma=Sigma_u, coverage=0.95,
                         stepdown=TRUE, simul=FALSE, R=1000, seed=101)
unis$margCSL <- CS_u$L
unis$margCSU <- CS_u$U

# plot - same pattern as Figure 1, scaled to 100 universities
ggplot(unis, aes(x = rank, y = reorder(name, rank))) +
  theme_bw() +
  geom_point(size = 0.7) +
  geom_errorbar(aes(xmin = margCSL, xmax = margCSU), width = 0.4) +
  scale_x_continuous(name = "Rank", limits = c(1,100),
                     breaks = c(1, seq(10,100,by=10))) +
  ylab(NULL) +
  labs(title = "Ranking of 100 Universities by Impact Factor",
       subtitle = "95% marginal confidence sets for ranks") +
  theme(axis.text.y = element_text(size = 5))
```

## Figure 2 — University Rankings

```{r figure2, fig.height=5.5, fig.width=9, echo=FALSE}
Sigma_u <- diag(unis$IFse^2)
CS_u <- csranks::csranks(unis$IFmean, Sigma=Sigma_u, coverage=0.95,
                         stepdown=TRUE, simul=FALSE, R=1000, seed=101)
unis$margCSL <- CS_u$L
unis$margCSU <- CS_u$U

ggplot(unis, aes(x = rank, y = reorder(name, rank))) +
  theme_bw() +
  geom_point(size = 0.7) +
  geom_errorbar(aes(xmin = margCSL, xmax = margCSU), width = 0.4) +
  scale_x_continuous(name = "Rank", limits = c(1,100),
                     breaks = c(1, seq(10,100,by=10))) +
  ylab(NULL) +
  labs(title = "Ranking of 100 Universities by Impact Factor",
       subtitle = "95% marginal confidence sets for ranks") +
  theme(
    axis.text.y = element_text(size = 5), 
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)) 
```

## Figure 2 — University Rankings (alt-view)

```{r figure2b, fig.height=5.5, fig.width=9, echo=FALSE}
# compute span and categories
unis$span <- unis$margCSU - unis$margCSL + 1
unis$span_cat <- cut(unis$span,
                     breaks = c(-Inf, 5, 10, 20, Inf),
                     labels = c("<=5","6-10","11-20",">20"))

# use full dataset (no top/bottom filter)
# order factor for plotting (so y-axis shows best first at top)
unis$name <- factor(unis$name, levels = unis$name[order(unis$rank)])

library(ggplot2)
ggplot(unis, aes(x = rank, y = reorder(name, rank))) +
  theme_bw() +
  geom_errorbar(aes(xmin = margCSL, xmax = margCSU, colour = span_cat),
                width = 0.4, size = 0.6, show.legend = TRUE) +
  geom_point(aes(colour = span_cat), size = 0.9, show.legend = TRUE) +
  scale_colour_manual(name = "Span (ranks)",
                      values = c("<=5"   = "red",
                                 "6-10"  = "black",
                                 "11-20" = "orange",
                                 ">20"   = "blue")) +
  scale_x_continuous(name = "Rank", limits = c(1, max(unis$rank, na.rm = TRUE)),
                     breaks = c(1, seq(10, max(unis$rank, na.rm = TRUE), by = 10))) +
  ylab(NULL) +
  labs(title = "Ranking of Universities by Impact Factor",
       subtitle = "95% marginal confidence sets for ranks (all ranks)") +
  theme(
    axis.text.y = element_text(size = 5),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right")
```

## Figure 2 — University Rankings (alt-view zoomed)

``` {r figure2c, fig.height=5.5, fig.width=9, echo=FALSE}
# compute span and categories
unis$span <- unis$margCSU - unis$margCSL + 1
unis$span_cat <- cut(unis$span,
                     breaks = c(-Inf, 5, 10, 20, Inf),
                     labels = c("<=5","6-10","11-20",">20"))

# filter to top 25 and bottom 25 (assumes rank 1 = best)
top_bottom <- subset(unis, rank <= 25 | rank > (max(rank, na.rm = TRUE) - 25))

# order factor for plotting (so y-axis shows best first at top)
top_bottom$name <- factor(top_bottom$name,
                          levels = top_bottom$name[order(top_bottom$rank)])

library(ggplot2)
ggplot(top_bottom, aes(x = rank, y = reorder(name, rank))) +
  theme_bw() +
  geom_errorbar(aes(xmin = margCSL, xmax = margCSU, colour = span_cat),
                width = 0.4, size = 0.6, show.legend = TRUE) +
  geom_point(aes(colour = span_cat), size = 0.9, show.legend = TRUE) +
  scale_colour_manual(name = "Span (ranks)",
                      values = c("<=5"   = "red",
                                 "6-10"  = "black",
                                 "11-20" = "orange",
                                 ">20"   = "blue")) +
  scale_x_continuous(name = "Rank", limits = c(1, max(top_bottom$rank, na.rm = TRUE)),
                     breaks = c(1, seq(10, max(top_bottom$rank, na.rm = TRUE), by = 10))) +
  ylab(NULL) +
  labs(title = "Ranking of Universities by Impact Factor",
       subtitle = "95% marginal confidence sets for ranks (top 25 and bottom 25)") +
  theme(
    axis.text.y = element_text(size = 6),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right")
```


## What the Figures Tell Us

Both figures show the same pattern the paper describes:

- Confidence sets are narrow at the very top and very bottom
- They widen in the middle of the distribution
- For journals, JEL is clearly ranked first but much of the middle cannot be
  separated from several adjacent ranks
- For universities, Chicago and Northwestern sit confidently at the top,
  but large clusters in the middle are statistically indistinguishable

This is the paper's central point visualized: **point rankings overstate
the precision of what we actually know.**


# Replication Notes

## What We Reproduced

Both main figures match the paper exactly. The confidence set bounds were
computed using the same `csranks` package the authors wrote, with identical
parameters and the same random seed (101) for reproducibility.

One technical issue we worked through: the current version of `csranks`
requires a covariance matrix rather than a vector of standard errors. The fix:

```{r show-fix, eval=FALSE}
# This fails with newer csranks:
csranks(jIF$IF, jIF$SE, ...)

# This works:
csranks(jIF$IF, Sigma = diag(jIF$SE^2), ...)
```

This reflects a deliberate API change by the authors to make the independence
assumption explicit rather than assumed silently.

## What Remains

- Additional figures from the companion paper as time permits
- full report of replication successes and failures
