---
title: "Statistical Uncertainty in the Ranking of Journals and Universities"
subtitle: "Replication of Mogstad, Romano, Shaikh & Wilhelm (2022)"
author: "Jason Jankoski, Alan Lamb, Linna Wang"
date: "ECON 687 — University of Maryland"
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE
)
knitr::opts_knit$set(root.dir = "../")
```

```{r install-packages, include=FALSE}
if (!requireNamespace("ggplot2",  quietly = TRUE)) install.packages("ggplot2")
if (!requireNamespace("dplyr",    quietly = TRUE)) install.packages("dplyr")
if (!requireNamespace("readxl",   quietly = TRUE)) install.packages("readxl")
if (!requireNamespace("devtools", quietly = TRUE)) install.packages("devtools")
if (!requireNamespace("csranks",  quietly = TRUE))
  devtools::install_github("danielwilhelm/R-CS-ranks")
```

```{r load-packages, include=FALSE}
library(ggplot2)
library(dplyr)
library(readxl)
library(csranks)
```


# The Paper

## What Question Does It Ask?

Rankings of journals and universities drive real decisions in hiring, promotion,
and funding. They are almost always reported as a single number with no
indication of how precisely that rank is estimated.

This paper asks: **given that impact factors are themselves estimated with
uncertainty, how much uncertainty surrounds the resulting ranks?**

The authors construct confidence sets for ranks. Rather than saying a journal
is ranked 5th, they report the full range of ranks that cannot be statistically
rejected at the 95% level.

## Method in Plain Terms

An impact factor is an average computed from a finite number of articles. It
carries a standard error. When you rank things by noisy estimates, the ranks
themselves are noisy.

The pipeline:

1. Start with point estimates (impact factors) and their standard errors
2. Build a covariance matrix capturing estimation uncertainty
3. Run a bootstrap simulation to generate possible rankings under that uncertainty
4. Report, for each journal or university, the set of ranks consistent with the data

The method is implemented in `csranks`, an R package written by the authors.

## What Did They Find?

**Journals** — 30 economics journals ranked by 2011 impact factor:

- 10 of the 30 cannot be statistically distinguished from the top 5
- Only 1 journal can be ruled out of the bottom 25

**Universities** — 100 universities ranked by average publication impact factor:

- Rankings at the very top and bottom are fairly precise
- Substantial uncertainty throughout the middle of the distribution

The main takeaway: point rankings overstate what the data can actually tell us.


# The Data

## Two Datasets, No External Access Required

Both datasets are included in the replication package.

| Dataset | Source | File |
|---|---|---|
| Journal impact factors | Stern (2013) | `data/journal IF/JEL_Data.xlsx` |
| University publications | Zimmermann (2013) | `data/Repec/top100unis.csv` |

David Stern assembled the journal citation data and shared it directly with the
authors. Christian Zimmermann maintains RePEC (Research Papers in Economics),
a database of economists' affiliations and publications, and shared the
underlying data.

## Journal Data — The Wrangling Problem

The raw Excel file is not a flat table. It has 233 sheets, one per journal.
Each sheet has rows for individual articles and columns for years, recording
how many citations each article received each year.

```{r show-import, eval=FALSE}
# You cannot just read_excel() this file and get something usable.
# The authors' convertdata.R loops through all 233 sheets,
# extracts the mean and SE from the summary rows at the bottom,
# and assembles a single clean data frame saved as IFdata.RData.

sheet_names <- excel_sheets(file.path("data", "journal IF", "JEL_Data.xlsx"))
length(sheet_names)  # 233 sheets
```

By the time we load `IFdata.RData`, the hard work is done. We get a clean
232-row data frame with columns: `journal`, `IF`, `SE`, `sigma`, `n`.

## Journal Data — Filtering to the KMS 30

The paper focuses on 30 journals pre-selected by Kalaitzidakis, Mamuneas &
Stengos (2003), an earlier study identifying leading outlets in economics.
We filter the 232-journal dataset to exactly those 30 by name.

```{r load-journal-data}
load(file.path("data", "journal IF", "IFdata.RData"))

top30names <- c("JEL", "QJE", "JEP", "JFE", "JPE", "Econometrica",
                "AER", "REStud", "RESTAT", "JLE", "JEEM", "JHumRes",
                "J INT ECON", "EconJ", "JMonEcon", "JEconometrics",
                "RAND J ECON", "J BUS ECON STAT", "JPubEcon",
                "JAppliedEconmet", "EUR ECON REV", "IER", "JET",
                "OXFORD B ECON STAT", "GAME ECON BEHAV",
                "J ECON DYN CONTROL", "SCAND J ECON",
                "ECONOMET THEOR", "Econ Theory", "EL")

jIF <- journalIF %>%
  filter(!is.na(IF), IF > 0, journal %in% top30names) %>%
  arrange(desc(IF)) %>%
  mutate(rank = row_number())

cat("Journals matched:", nrow(jIF), "of 30")
```

## Journal Data — Descriptive Statistics

```{r journal-summary}
jall <- journalIF %>% filter(!is.na(IF), IF > 0)
cat("--- All 231 Valid Journals ---\n")
cat("Mean IF:  ", round(mean(jall$IF), 4), "\n")
cat("Median IF:", round(median(jall$IF), 4), "\n")
cat("SD of IF: ", round(sd(jall$IF), 4), "\n")
cat("Range:    [", round(min(jall$IF), 4), ",", round(max(jall$IF), 4), "]\n\n")

cat("--- KMS Top 30 ---\n")
cat("Mean IF:  ", round(mean(jIF$IF), 4), "\n")
cat("Median IF:", round(median(jIF$IF), 4), "\n")
cat("SD of IF: ", round(sd(jIF$IF), 4), "\n")
cat("Mean SE:  ", round(mean(jIF$SE), 4), "\n")
cat("Range:    [", round(min(jIF$IF), 4), ",", round(max(jIF$IF), 4), "]\n")
```

## Journal Data — What We Are Working With

```{r journal-table}
jIF %>%
  select(Rank = rank, Journal = journal, IF, SE, N_articles = n) %>%
  mutate(IF = round(IF, 3), SE = round(SE, 3)) %>%
  head(10) %>%
  knitr::kable(caption = "Top 10 journals by 2011 impact factor")
```

## University Data — Already Clean

The university data required less wrangling. The authors' `cleandata.R` and
`keeptop100.R` scripts processed the raw RePEC text file upstream. By the time
we load `top100unis.csv` it is ready to use: 100 universities with average
impact factors and standard errors already computed.

```{r load-university-data}
unis <- read.csv(file.path("data", "Repec", "top100unis.csv")) %>%
  filter(!is.na(IFmean)) %>%
  arrange(desc(IFmean)) %>%
  mutate(rank = row_number())

# IFmean = average IF of publications by authors affiliated with that university
# IFse   = standard error of that average
# n      = number of affiliated authors in RePEC

cat("Universities:", nrow(unis), "| Total authors:", sum(unis$n))
```

## University Data — Descriptive Statistics

```{r university-summary}
cat("Mean IFmean:  ", round(mean(unis$IFmean), 4), "\n")
cat("Median IFmean:", round(median(unis$IFmean), 4), "\n")
cat("SD of IFmean: ", round(sd(unis$IFmean), 4), "\n")
cat("Mean IFse:    ", round(mean(unis$IFse), 4), "\n")
cat("IF range:     [", round(min(unis$IFmean), 4), ",",
                       round(max(unis$IFmean), 4), "]\n")
```

```{r university-table}
unis %>%
  slice(1:10) %>%
  select(Rank = rank, University = name, IFmean, IFse, N_authors = n) %>%
  mutate(IFmean = round(IFmean, 3), IFse = round(IFse, 3)) %>%
  knitr::kable(caption = "Top 10 universities by average impact factor")
```


# The Code

## The Full Pipeline

```
data/journal IF/convertdata.R     # Excel (233 sheets) -> IFdata.RData
data/journal IF/journalranking.R  # confidence sets -> Figure 1
data/Repec/cleandata.R            # raw RePEC text -> cleaned data
data/Repec/keeptop100.R           # filter to top 100 universities
data/Repec/emp.topunis.R          # confidence sets -> Figure 2
```

Our `code/00_data_check.R` runs the key steps independently to verify
the pipeline without running the full `master.R`.

The four-step logic repeated for both datasets:

> **Load clean data, filter to sample, run bootstrap, plot results**

## The Key Statistical Step — csranks

```{r show-csranks, eval=FALSE}
# Step 1: build a covariance matrix from the standard errors
# diag(SE^2) puts variances on the diagonal, zeros off-diagonal
# This assumes estimation errors are independent across journals
Sigma <- diag(jIF$SE^2)

# Step 2: run the bootstrap
CS <- csranks::csranks(
  jIF$IF,           # point estimates
  Sigma = Sigma,    # covariance matrix
  coverage = 0.95,  # 95% confidence level
  stepdown = TRUE,  # iterative procedure, gives tighter sets
  simul    = FALSE, # marginal sets (each journal covered individually)
  R        = 1000,  # bootstrap draws
  seed     = 101    # reproducibility
)

# Step 3: attach bounds to the data frame
jIF$CSL <- CS$L   # lower bound of rank confidence set
jIF$CSU <- CS$U   # upper bound of rank confidence set
```

## From Bootstrap Output to Figure

Once `csranks()` returns the bounds, the figure follows directly.
The bounds become the error bar limits in ggplot.

```{r show-plot, eval=FALSE}
ggplot(jIF, aes(x = rank, y = reorder(journal, rank))) +
  theme_bw() +
  geom_point(size = 2) +             # dot = point rank estimate
  geom_errorbar(                      # bar = confidence set for rank
    aes(xmin = CSL, xmax = CSU),
    width = 0.4
  ) +
  scale_x_continuous(
    name = "Rank",
    limits = c(1, 30),
    breaks = c(1, 5, 10, 15, 20, 25, 30)
  ) +
  ylab(NULL)
```

The x-axis is rank, not impact factor. The horizontal bars show the range of
ranks the data cannot rule out for each journal at the 95% level.

## Marginal vs. Simultaneous

`simul = FALSE` produces **marginal** confidence sets. Each journal's set
covers its true rank with 95% probability on its own.

`simul = TRUE` produces **simultaneous** sets. All true ranks are covered
jointly at 95%. These are wider because the guarantee is stronger.

The paper shows both. Figure 1 uses marginal sets. The appendix shows
simultaneous sets for comparison. We replicate both.


# The Figures

## Figure 1 — Journal Rankings

```{r figure1, fig.height=4.5, fig.width=9, echo=FALSE}
cs_path <- file.path("data", "journal IF", "journalIFrankings.RData")
if (file.exists(cs_path)) {
  load(cs_path)
  plot_data <- journalIF.top30
} else {
  Sigma <- diag(jIF$SE^2)
  CS <- csranks::csranks(jIF$IF, Sigma=Sigma, coverage=0.95,
                          stepdown=TRUE, simul=FALSE, R=1000, seed=101)
  plot_data <- jIF %>% mutate(margCSL = CS$L, margCSU = CS$U)
}

ggplot(plot_data, aes(x = rank, y = reorder(journal, rank))) +
  theme_bw() +
  geom_point(size = 2) +
  geom_errorbar(aes(xmin = margCSL, xmax = margCSU), width = 0.4) +
  scale_x_continuous(name = "Rank", limits = c(1,30),
                     breaks = c(1, 5, 10, 15, 20, 25, 30)) +
  ylab(NULL) +
  labs(title = "Ranking of 30 Journals by 2011 Impact Factor",
       subtitle = "95% marginal confidence sets for ranks")
```

## {.flexbox .vcenter}

```{r figure2, fig.height=6.5, fig.width=8, echo=FALSE}
Sigma_u <- diag(unis$IFse^2)
CS_u <- csranks::csranks(unis$IFmean, Sigma=Sigma_u, coverage=0.95,
                          stepdown=TRUE, simul=FALSE, R=1000, seed=101)
unis$margCSL <- CS_u$L
unis$margCSU <- CS_u$U

ggplot(unis, aes(x = rank, y = reorder(name, rank))) +
  theme_bw() +
  geom_point(size = 0.7) +
  geom_errorbar(aes(xmin = margCSL, xmax = margCSU), width = 0.4) +
  scale_x_continuous(name = "Rank", limits = c(1,100),
                     breaks = c(1, seq(10,100,by=10))) +
  ylab(NULL) +
  labs(title = "Ranking of 100 Universities by Impact Factor",
       subtitle = "95% marginal confidence sets for ranks") +
  theme(axis.text.y = element_text(size = 5))
```

## What the Figures Tell Us

Both figures show the same pattern the paper describes:

- Confidence sets are narrow at the very top and very bottom
- They widen substantially in the middle of the distribution
- For journals, JEL is clearly ranked first but much of the middle cannot be
  separated from several adjacent ranks
- For universities, Chicago and Northwestern sit confidently at the top,
  but large clusters in the middle are statistically indistinguishable

This is the paper's central point visualized: **point rankings overstate
the precision of what we actually know.**


# Replication Notes

## What We Reproduced

Both main figures match the paper. The confidence set bounds were computed
using the same `csranks` package the authors wrote, with identical parameters
and the same random seed (101) for reproducibility.

One technical issue we worked through: the current version of `csranks`
requires a covariance matrix rather than a vector of standard errors. The fix:

```{r show-fix, eval=FALSE}
# This fails with newer csranks:
csranks(jIF$IF, jIF$SE, ...)

# This works:
csranks(jIF$IF, Sigma = diag(jIF$SE^2), ...)
```

This reflects a deliberate API change by the authors to make the independence
assumption explicit rather than assumed silently.

## What Remains for the Final Report

- Written discussion of the econometric foundations of the confidence set method
- Side-by-side comparison of our figures with the paper's originals
- Walkthrough of the RePEC cleaning scripts (`cleandata.R`, `keeptop100.R`)
- Discussion of replication success and any remaining differences
